---
title: "General Information"
author: "Olabiyi Obayomi"
date: "4 5 2021"
output: html_document
params:
  metadata_file: 'mapping_files/metadata.tsv'
  group: "Biofilter_media"
  analysis: "kaiju"
---

<style type="text/css">
.main-container {
max-width: 1500px;
margin-left: auto;
margin-right: auto;
}
</style>


```{r echo=FALSE, include=FALSE, comment='', message=FALSE, warning = FALSE}

# Raw feature table file outputs from running kaiju2table in linux
files <- list.files(here::here('data/kaiju/'),pattern = '.tsv', full.names = TRUE) 
names(files) <- files %>% str_replace_all(pattern = '.+\\.(\\w+)\\.tsv',replacement = '\\1')  %>% toTitleCase()
analysis_name <- "colored_short_kaiju_taxonomy"
taxon_levels <- c("Phylum", "Class",  "Order", "Family", "Genus", "Species")
names(taxon_levels) <- taxon_levels
metadata_file <- here::here('mapping_files/metadata.tsv')
plot_dir <- here::here(glue::glue('results/kaiju/'))
# Groups used for testing interactions and for betaDiversity ploting
main_groups <- c("Biofilter_media", "Biofilter_configuration", "Event") 
# Groups used for alpha diversity statistics, box plots and barz plots
groups <- c("Biofilter_media", "Group")
names(groups) <- groups
x_labs <- c("Biofilter Media", "Group")
names(x_labs) <- x_labs

Biofilter_media_colors <- c(G="chocolate4", S="darkgoldenrod1")

Biofilter_configuration_colors <- c(CAR="mediumseagreen", NV="red2", NVNB="mediumblue")

Event_colors <- c(A="lightsalmon1", B="plum2",  C="lightslateblue", D="mislvrose3", E="lightseagreen")

Group_colors <- c(Negative_control="white",Slurry_control="brown",Positive_control="tan4",
                  A.G_CAR="seagreen1", B.G_CAR="seagreen2", 
                  C.G_CAR="seagreen3", D.G_CAR="seagreen4", E.G_CAR="palegreen4",
                  A.G_NV="chocolate1", B.G_NV="chocolate2",
                  C.G_NV="chocolate3", D.G_NV="chocolate4", E.G_NV="coral4",
                  A.G_NVNB="bisque2", B.G_NVNB="bisque3", C.G_NVNB="bisque4",
                  D.G_NVNB="azure2", E.G_NVNB="azure3",
                  A.S_CAR="chartreuse", B.S_CAR="chartreuse3", 
                  C.S_CAR="chartreuse4", D.S_CAR="darkolivegreen1", E.S_CAR="darkgreen",
                  A.S_NV="darkorange", B.S_NV="darkorange3", 
                  C.S_NV="darkorange4", D.S_NV="orange", E.S_NV="orange3",
                  A.S_NVNB="gray", B.S_NVNB="darkslategray1", C.S_NVNB="darkslategray4",
                  D.S_NVNB="darkslategrey", E.S_NVNB="gray7")

annotation_colors <- list(Biofilter_media=c(NONE='white', G="chocolate4", S="darkgoldenrod1"),
                          Biofilter_configuration=c(NONE="white",CAR="mediumseagreen", NV="red2", NVNB="mediumblue"),
                          Event=c(Negative_control="powderblue",Slurry_control="maroon1",Positive_control="black",
                                  A="lightsalmon1", B="plum2",  C="lightslateblue", D="mislvrose3", E="lightseagreen"))


sample_order <- 
  c('59.EX-B_17052019', '60.B-SC_22052019', '61.TC1_C1_30092019',
    
    '01.G-CAR-11_200619_G2', '11.G-CAR-14_29082019','12.G-CAR-7_29082019', '23.G-CAR-3_12112019',
    '24.G-CAR-7_12112019', '35.G-CAR-11_10122019', '36.G-CAR-7_10122019', '47.G-CAR-3_10012020', '48.G-CAR-7_10012020',
    
    '02.G-NV-2_20062019', '03.G-NV-8_20062019', '13.G-NV-2_29082019', '14.G-NV-8_290819', '25.G-NV-2_12112019', '26.G-NV-8_12112019', '37.G-NV-2_10122019', '38.G-NV-8_10122019', '49.G-NV-2_10012020', '50.G-NV-8_10012020',
    
    '04.G-NVNB-12_20062019', '05.G-NVNB-4_20062019', '15.G-NVNB-12_29082019', '16.G-NVNB-6_29082019', '27.G-NVNB-12_12112019', '28.G-NVNB-4_12112019', '39.G-NVNB-12_10122019', '40.G-NVNB-6_10122019', '51.G-NVNB-12_10012020', '52.G-NVNB-4_10012020',
    
    '06.S-CAR-19_20062019', '07.S-CAR-24_20062019', '17.S-CAR-19_29082019', '18.S-CAR-31_29082019',
    '29.S-CAR-19_12112019', '30.S-CAR-31_12112019', '41.S-CAR-19_10122019', '42.S-CAR-24_10122019', '53.S-CAR-24_10012020', '54.S-CAR-31_10012020',
    
    '08.S-NV-26_20062019', '09.S-NV-30_20062019', '19.S-NV-18_29082019', '20.S-NV-26_29082019', '31.S-NV-18_12112019', '32.S-NV-26_12112019', '43.S-NV-26_10122019', '44.S-NV-30_10122019', '55.S-NV-26_10012020', '56.S-NV-30_10012020',
    
    '10.S-NVNB-20_20062019', '21.S-NVNB-20_29082019','22.S-NVNB-32_29082019', '33.S-NVNB-20_12112019', '34.S-NVNB-28_12112019', '45.S-NVNB-20_10122019', '46.S-NVNB-28_10122019', '57.S-NVNB-20_10012020', '58.S-NVNB-28_10012020'
  )



Group_order <- c('Negative_control','Slurry_control','Positive_control',
                 "A.G_CAR", "B.G_CAR", "C.G_CAR", "D.G_CAR", "E.G_CAR",
                 "A.G_NV", "B.G_NV", "C.G_NV", "D.G_NV", "E.G_NV",
                 "A.G_NVNB", "B.G_NVNB", "C.G_NVNB", "D.G_NVNB", "E.G_NVNB",
                 "A.S_CAR", "B.S_CAR", "C.S_CAR", "D.S_CAR", "E.S_CAR",
                 "A.S_NV", "B.S_NV", "C.S_NV", "D.S_NV", "E.S_NV",
                 "A.S_NVNB", "B.S_NVNB", "C.S_NVNB", "D.S_NVNB", "E.S_NVNB")

Biofilter_media_order <- c("NONE", "G", "S")

Event_order <- c('Negative_control','Slurry_control','Positive_control', 'A', 'B', 'C', 'D', 'E')

Biofilter_configuration_order <- c("NONE", "CAR", "NV", "NVNB")




factors_order <-  list("Sample_name" = sample_order ,
                       "Group" = Group_order,
                       "Biofilter_media" = Biofilter_media_order,
                       "Event" = Event_order,
                       "Biofilter_configuration" = Biofilter_configuration_order)

```

```{r echo=FALSE, include=FALSE, comment='', message=FALSE, warning = FALSE}

# Prepare feature tables
taxon_tables <- map(.x = files,.f = process_kaiju_table)
# sort the taxon tables by taxon level from phylum to species
taxon_tables <- taxon_tables[taxon_levels]
```


# General information


This report is a short version of my initial report. In this report I only report results for "Group" comparisons. However, the statistics for the Group (combination of all factors at different time points) is not statistically sound and hence unreliable because each time point had less than three samples necessary for proper statistical analysis. All the tables and many plots presented in this report are interactive, searchable and downloadable.


### Sample metadata table


The table below shows the metadata used for the analyses.


```{r echo=FALSE, include=TRUE, comment='', message=FALSE, warning = FALSE}
# Prepare the metadata
# Read-in the metatdata

columns2keep <- c("Sample_name", "Plate", "Well",  "Group" , 
                  "Event" , "Date", "Biofilter_media", 
                  "Biofilter_configuration")

metadata <-  read_delim(file = metadata_file, delim = "\t", col_names = TRUE) %>%
  as.data.frame(stringAsFactor=FALSE)

# Correct the name of "61.TC1 C1_30092019" to "61_TC1 C1_30092019" to match that used in the feature table
metadata$Sample_name <-  gsub(pattern = " ",replacement = "_", x = metadata$Sample_name)

metadata$Date <- lubridate::dmy(metadata$Date)
rownames(metadata) <- metadata$Sample_name


for(column in colnames(metadata)){
  if(is.character(metadata[,column])){
    metadata[,column]<- tidyr::replace_na(metadata[,column], 'NONE')
    metadata[,column] <- gsub(pattern = "NA ",replacement = '', x = metadata[,column])
  }
}


# Set factor order
for(factor_column in names(factors_order)){
  
  metadata[,factor_column] <- factor(x = metadata[,factor_column], levels = factors_order[[factor_column]])
  
}


create_dt(metadata)
metadata <- metadata[,columns2keep]
metadata <- metadata %>% rownames_to_column(var = "Sample.ID") %>% as.data.frame()

rownames(metadata) <- metadata[,"Sample.ID"]

# remove control samples
no_control_metadata <- droplevels(metadata[-(59:61),])
```


### The number of samples per treatment category

This table shows the count of how many samples (N) were available at each category level. 



```{r echo=FALSE, include=TRUE, comment='', message=FALSE, warning = FALSE}
categories_freq_df <- metadata %>% 
  mutate(Sample_name=as.character(Sample_name)) %>% 
  summary(maxsum=100) %>% 
  as.data.frame() 

categories_freq_df[complete.cases(categories_freq_df),] %>% 
  dplyr::select(Var2:Freq) %>% 
  dplyr::filter(str_detect(Var2,"Bio|Event|Group")) %>% 
  dplyr::select(Category=Var2, N=Freq) %>% 
  droplevels() %>% 
  separate(col = N,into = c("Treatment","N"), sep = ":") %>% 
  dplyr::mutate(N=as.numeric(N)) %>% 
  dplyr::filter(N > 0) %>% 
  create_dt()




```


```{r echo=FALSE, include=FALSE, comment='', message=FALSE, warning = FALSE}
save.image(file = glue::glue('{plot_dir}/{analysis_name}.RData'))
sessionInfo()
```
